---
title: "NYPD Frisk"
output:
  html_document: default
  html_notebook: default
---
#### Loading R packages
```{r}
library(lubridate)
```
#### Loading NYPD Dataset
```{r}
nypd = read.csv("C:/Users/Supriya Khadake/Documents/Projects/NYPD_Frisk/nypd_frisk_data.csv")
```

###Data Exploration

#### Keeping relevant attributes only
```{r}
nypd = subset(nypd, select = -c(year,ser_num,recstat,trhsloc,crimsusp,arstmade,sumoffen,compyear,comppct,perobs,perstop,adtlrept,adtlrept,pistol,riflshot,asltweap,knifcuti,machgun,othrweap,pf_hands,pf_wall,pf_grnd,pf_drwep,pf_ptwep,pf_baton,pf_hcuff,pf_pepsp,pf_other,radio,ac_rept,ac_inves,sb_hdobj,sb_outln,sb_admis,sb_other,repcmd,revcmd,cs_bulge,cs_casng,cs_cloth,cs_descr,cs_drgtr,cs_furtv,cs_lkout,cs_objcs,cs_other,cs_vcrim,offverb,offshld,forceuse,dob,addrtyp,rescode,premtype,premname,addrnum,stname,stinter,crossst,aptnum,state,zip,addrpct,sector,beat,post,dettypCM,lineCM,detailCM,xcoord,ycoord,officrid,arstoffn,sumissue,explnstp))

colnames(nypd)
```

```{r}
nypd$frisked = nypd$frisked
nypd$pct = nypd$pct
nypddatestop = nypd$datestop
nypd$timestop = nypd$timestop
nypd$inout = nypd$inout
nypd$typeofid = nypd$typeofid
nypd$othpers = nypd$othpers
nypd$offunif= nypd$offunif
nypd$searched = nypd$searched
nypd$contrabn = nypd$contrabn
nypd$rf_vcrim = nypd$rf_vcrim
nypd$rf_othsw = nypd$rf_othsw
nypd$ac_proxm = nypd$ac_proxm
nypd$rf_attir = nypd$rf_attir
nypd$rf_vcact = nypd$rf_vcact
nypd$ac_evasv = nypd$ac_evasv
nypd$ac_assoc =	nypd$ac_assoc
nypd$rf_rfcmp =	nypd$rf_rfcmp
nypd$ac_cgdir =	nypd$ac_cgdir
nypd$rf_verbl =	nypd$rf_verbl
nypd$ac_incid =	nypd$ac_incid
nypd$ac_time =	nypd$ac_time
nypd$rf_knowl =	nypd$rf_knowl
nypd$ac_stsnd =	nypd$ac_stsnd
nypd$ac_other =	nypd$ac_other
nypd$rf_furt =	nypd$rf_furt
nypd$rf_bulg =	nypd$rf_bulg
nypd$sex =	nypd$sex
nypd$race =	nypd$race
nypd$age =	nypd$age
nypd$weight =	nypd$weight
nypd$haircolr =	nypd$haircolr
nypd$eyecolor =	nypd$eyecolor
nypd$build =	nypd$build
nypd$othfeatr =	nypd$othfeatr
nypd$city =	nypd$city
nypd$height =	nypd$height
```

```{r}
head(nypd)
tail(nypd)
```

#### Descriptive statistics of the dataset before data transformaion
```{r}
summary(nypd)
```

#### Structure of the dataset before data transformation
```{r}
str(nypd)
```

### Data Transformation and Cleaning

#### Aggregated variables like height in feet and inches to one vriables height
```{r}
nypd$height = nypd$ht_feet+(nypd$ht_inch*0.0833333)
nypd$height = as.integer(round(nypd$height))
nypd = subset(nypd, select = -c(ht_feet, ht_inch))
```

####  Extracting month from date
```{r}
nypd$datestop = mdy(nypd$datestop)
nypd$datestop = as.factor(month(as.POSIXlt(nypd$datestop, format="%d/%m/%Y")))
```

#### Conveting timestop to factors
```{r}
nypd$timestop = as.factor(ifelse(nypd$timestop > 0 & nypd$timestop < 559, 1, ifelse(nypd$timestop > 600 & nypd$timestop < 1159, 2, ifelse(nypd$timestop > 1200 & nypd$timestop < 1759, 3,4 ))))
```

#### Replacing age and weight irrelevant values with mean
#### Converting typeofid to binary formaat if R(refused) then 0 else O,P,R(showed) = 1
#### Was stopped inside or outside values to binary i.e. if O then 0 else 1
```{r}
nypd$age = ifelse(nypd$age > 100 | nypd$age < 0, 28.96, nypd$age)
nypd$weight = as.integer(ifelse(nypd$weight > 500 | nypd$weight <20, 171.4, nypd$weight))
nypd$typeofid = as.factor(ifelse(nypd$typeofid == 'R', 0, 1))
nypd$inout = as.factor(ifelse(nypd$inout == 'O', 0, 1))
```

#### Converting datatype of datestop & timestop from num to int
```{r}
nypd$datestop = as.integer(nypd$datestop)
nypd$timestop = as.integer(nypd$timestop)
```


#### Descriptive statistics of the dataset after data transformation
```{r}
summary(nypd)
```

#### Structure of the dataset
```{r}
str(nypd)
```

```{r}
head(nypd)
tail(nypd)
```

```{r fig.width = 14, fig.height = 10}
correlation_var = nypd[c(1:3,6:27,30,37)]
correlation = cor(correlation_var)
corrplot(correlation)
```
```{r}
colnames(nypd)
```

```{r}


```

```{r}
cor(cbind(pct, frisked))
```
































